---
title: "Homework 11"
author: "Xiangyu Ren"
output:
  pdf_document: default
  word_document: default
---

1. Consider an experiment to determine whether or not having an automatic transmission affects gas mileage in cars.  It is also known that the horsepower of a car affects gas mileage, so the horsepower of each car used is also recorded.  The following R code will read in the data to R
```{r, warning=F, message=F}
library(lsmeans)
library(car)
library(multcompView)
library(lme4)
library(lmerTest)
library(jtools)
library(interactions)
```

```{r}
cars <- mtcars[,c("am","mpg","hp")]
head(cars)
## type "cars" into R to see the full data
cars$am = as.factor(cars$am)
cars
```

Fit an appropriate model to this data, and interpret the results.

$$Y_{it}=\mu+\alpha_i+\beta* x_j+(\alpha\beta)_i* x_{it}+\epsilon_{ijt},\quad \epsilon_{it} \stackrel{iid}{\sim} N(0,\sigma^2)$$
```{r}
fit.ancova = aov(mpg ~ am + hp + am : hp, data = cars)
plot(fit.ancova)
anova(fit.ancova)
```
Hypothesis test with $\alpha=0.05$

'am:hp' line tests with

$$H_0:\;\;(\alpha\beta)_i=0\text{ for all }i\;\;vs.\;\;H_a:\;\;\text{at least one of the intercept terms }(\alpha\beta)_i\ne0$$

'hp' line tests with

$$H_0:\;\;\beta=0\;\;vs.\;\;H_a:\;\;\beta\ne0$$

By the table, we can see that the p-value of am:hp is greater than $\alpha$, so we fail to reject the null, and conclude that the interaction between am and hp is not significant; the p-values of either am or hp are less than $\alpha$, so these two variables are significant.

```{r}
interact_plot(fit.ancova, pred ="hp", modx="am")
lsmeans(fit.ancova, ~ hp)
```
$\beta=20.6$ If cars have same hp, then cars with $am=1$ have higher mpg than cars with $am=0$. If cars have same am, then mpg decreases as hp decreases.

2. The "high school and beyond" survey looked at differences in socioeconomic status, race, gender, and types of schools and students' SAT scores.  The data can be read into R using the following code:

The goal of this study is to see if socio-economic status (ses) and school type (schtyp) affect students math scores, after accounting for how well they do on their writing score.  Student's writing score should be treated as a continuous covariate.  Conduct a full analysis of this data.  Fit an appropriate model, and interpret the results.

```{r}
hsb2= read.table("hsb2.csv")## from https://stats.idre.ucla.edu/stat/data/hsb2.csv
head(hsb2)
## type "hsb2" into R to see the full data
hsb2$ses = as.factor(hsb2$ses); hsb2$schtyp = as.factor(hsb2$schtyp)
hsb2
```
$$Y_{ijkt} = \mu+\alpha_i+\gamma_k+(\alpha\gamma)_{ik}+\beta*x_j+(\alpha\beta)_{i}*x_{it}+(\gamma\beta)_k*x_{kt}+\epsilon_{ijkt},\quad \epsilon_{ijt} \stackrel{iid}{\sim} N(0,\sigma^2)$$
```{r}
fit.ancova = aov(math ~ ses + schtyp + write + write : ses + write : schtyp + ses : schtyp, data = hsb2)
plot(fit.ancova)
```
By the graphs, it appears to be both non-constant variance and non-normality in the residuals under the untransformed data, doesn't satisfy the assumption of ANCOVA. 
```{r}
anova(fit.ancova)
```
By the table, p-values of interactions of ses and write, and schtyp and write are larger than $\alpha$, so fail to reject the null and conclude that they are not significant, p-values for the rest:ses, schtyp, write, interaction between ses adn schtyp are smaller than $\alpha$, thus, they are significant.
```{r}
interact_plot(fit.ancova, pred = "write",modx = "ses")
lsmeans(fit.ancova, ~ write)
```
$\beta=3.94$ If students have same write, then math of students with $ses=3>ses=2>ses=1$. If students have same ses, then math increases as write increases.
