---
title: "Homework X"
author: "Put your name here"
output: html_notebook
---

### Question 1: Cars is an in-built dataset in R. Print the first couple of observations in the dataset

```{r}
#R code for the question 1 goes here
head(cars) #head() function print the first couple of rows of a dataset
```

### Question 2: What is the dimension of the dataset? 
```{r}
dim(cars)
```
### Question 3: How many observations and how many variables are there in the Cars dataset? 
There are two variables: Speed and Distance and 50 observations in the *cars* dataset.

### Question 4: Make a scatter plot of the data and comment on your observations

```{r}
plot(cars$speed,cars$dist,xlab="Speed", ylab= "Stopping Distance")
```
From the above plot we can see that there is a positive relationship between the speed and the stopping distance, perhaps the relationship is somewhat non-linear. 

### Question 5: Obtain the least squares regression line for predicting stopping distance from speed. 

```{r}
carsFit=lm(dist~speed, data=cars)
summary(carsFit)
```
### Question 6: Write the equation of the least squares regression line. 
When you type up an equation you can your equation within double dollar signs like 
$$\hat{y}=\hat{\beta}_0+\hat{\beta}_1x$$
For this particular example, equation of the least squares regression line is
$$\hat{y}=-17.58+3.93x$$

### Question 7: Say you want to type up the equation of a confidence interval. Then the general equation for the confidence interval for mean is
$$\bar{x}\pm t_{\frac{\alpha}{2},(n-1)}\frac{s}{\sqrt{n}}$$
Suppose for a particular example the sample average is 278.6, sample standard deviation is 53.9, sample size is 30 and t-multiplier is 2.045. Then you can plug-in the values to the formula as below to show your work:
$$278.6\pm2.045\times\frac{53.9}{\sqrt{30}}=(258.5,298.7)$$

### Question 8: Let's take the hypothesis testing question we did in lesson 0: "Suppose the faculty member has a realtor and he cliams average house price is \$ 225,000. Faculty member wants to see if it is plausible. This is how you would show the 6-steps:

1. Step 1: $H_0:\mu \geq 35$ Vs. $H_1: \mu \neq 35$
2. Step 2: We already plot the data in class. Histogram of the data looked roughly symmetric. No evidnce of violations in normality. Choose$\alpha = 5\% $
3. Step 3: $t=\frac{(278.6-255)}{53.9/\sqrt{30}}=2.398$
4. Step 4:P value or the rejection region. I will show both here:
  + P value approach:$2P(T>2.398)=0.02314389$. If you bracket the p value using the t-table, it would be between 0.02 and 0.05
  + Rejection Region approach: Reject if $t\leq -2.045$ or $t\geq 2.045$
5. Step 5: Reject H0 as P value is less than 0.05 (test statistic falls in the rejection region)
6. Step 6: At 5% significance level, we have sufficient evidence to conclude that realtors'claim about average house price in State College is wrong. 


### Question 9: Follow the below syntax to create a table to show your work for variable selection. 

Variable(s) | $R^2$ | Adj. $R^2$
------------------------------------------------------------------------------ | ------------- | -------------
RottenTomatoes | "include the R2value" | "include the adj R2value" 
AudienceScore |  | 
TheatersOpenWeek |  | 
BOAverageOpenWeek |  | 
DomesticGross |  | 
DomesticGross, RottenTomatoes |  | 
DomesticGross, AudienceScore | | 
DomesticGross, TheatersOpenWeek |  | 
DomesticGross, BOAverageOpenWeek |  | 
DomesticGross, RottenTomatoes, AudienceScore |  | 
DomesticGross, RottenTomatoes, TheatersOpenWeek |  | 
DomesticGross, RottenTomatoes, BOAverageOpenWeek |  | 
DomesticGross, RottenTomatoes, AudienceScore, TheatersOpenWeek |  | 
DomesticGross, RottenTomatoes, AudienceScore, BOAverageOpenWeek |  | 


